### 2016-06-18 sections/intro0.0.tex

%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%Intro
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------
\section{Introduction}
%General information about galaxies %(Maybe it is too generic?!!)
Inside each galaxies are millions of stars and large amount of dust and gas, which all have affect on each others' formation, evolution and extinction.
Galaxies are an active and complicated systems, that are constantly changing.
There are extensive amount of studies about correlations between galaxies' emission and their physical properties, both observationally and theoretically %(Some references here). 
With all these studies our knowledge about the underlying processes in formation and evolution of galaxies expands every day, but the details of how these process work are still undetermined.

% Nearby galaxies and their importance 
Nearby galaxies plays an important role in improving studies about galaxies' formation and evolution.
These galaxies are close enough to be able to observe various regions inside the galaxies in details.
As a result, many studies were devoted to find relations between physical properties of the galaxies in both spatially resolved regions and galaxies as a whole %(Some more references here)
, and many others used these results to analyze data from high-redshift galaxies.%(Some more references here)
The amount of the data that is acquired from nearby galaxies is increasing daily.
Advances in technology helps to both observe and store enormous amount of data from the sky.
Telescopes and surveys gatherer data from nearby galaxies in different wavelengths and from these data we can measure physical properties of galaxies (e.g. stellar mass, star formation rate (SFR), dust mass, gas mass). %(Some more references here)

M31, with distance of $\sim 0.78$~Mpc~\citep{McConnachie05}, is one of the closest spiral galaxies to the Milky way and provide us very unique view of inside of a spiral galaxy.
Therefore,

% A little bit about PAH and other properties of nearby galaxies?!

% data mining and clustering in general; when we have so many data and we want to map them
\cite{Ball10} did an extensive review of data mining and machine learning and their usage in astronomy.

%SOM is a good method for knowledge discovery projects;
%SOM in Astronomy and mention we do not have any SOM in nearby galaxies
%this project 1 M31 and its extensive data

%supervised and unsupervised network